---
title: "Cleaning baseline isotope data"
author: "Camila Vargas"
date: "10/12/2021"
output: html_document
---

Cleaning baseline isotope data provided by Ana Miller-ter Kuile to PADL. Raw data [here.](https://github.com/anamtk/palmyra/tree/main/raw_data/isotope_data)



## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)

## Sets R not to use scientific notations
options(scipen=999) 

```


## Check all necesary files are in the raw_data folder

```{r}
## File path to raw data used in this script
raw_data_path <- here::here("raw_data")

# Count files inside the raw_data folder to make sure the number of files downloaded is what is expected.
length(list.files(raw_data_path, pattern = "baseline"))

## in this case there should be **2** files
```


## Table with all the raw data files
```{r}

## List with all xls files
all_xls <- tibble(list.files(raw_data_path, pattern = "baseline")) %>% #c("xls", "xlsx")
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         ### UPDATE
         type = "baseline") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)

```


## Read data

```{r}

## 2008 Algae

## check sheet names
excel_sheets(all_xls$path[1]) ## Only one sheet

baseline_algae <- read_excel(all_xls$path[1], sheet = "Sheet1", skip = 4) %>% 
  clean_names()

## 2008 Guano
excel_sheets(all_xls$path[2]) ## Only one sheet

baseline_guano <- read_excel(all_xls$path[4], sheet = "Sheet1", skip = 4) %>% 
  clean_names()

```

Both datasets have the same colnames! But ID#3 in Algae data is Location = Palmyra and in Guano is Type = Guano.

Don't mess up with colnames. We will just keep names as id_1, id_2 and let the author describe what each ide means!


## Cleaning baseline data

```{r}
baseline_complete <- bind_rows(baseline_algae, baseline_guano) %>% 
  rename(amp_v_n = amp_v_6,
         amp_v_c = amp_v_7,
         d15n_air = d15n,
         d13c_pdb = d13c,
         c_n_atomic = c_n) %>% 
  filter(!is.na(date))
  
```


## Save to clean data

```{r}
write_csv(baseline_complete, here::here("clean_data/2009_palmyra_baseline_isotopes.csv"))
```



