---
title: "Cleaning animal isotopes data"
author: "Camila Vargas"
date: "10/12/2021"
output: html_document
---

Cleaning and formating other animal isotope data provided by Ana Miller-ter Kuile to PADL. Raw data [here.](https://github.com/anamtk/palmyra/tree/main/raw_data/isotope_data)

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)

## Sets R not to use scientific notations
options(scipen=999) 

```


## Check all necesary files are in the raw_data folder

```{r}
## File path to raw data used in this script
raw_data_path <- here::here("raw_data")

# Count files inside the raw_data folder to make sure the number of files downloaded is what is expected.
length(list.files(raw_data_path, pattern = "animals"))

## in this case there should be **6** files
```


## Table with all the raw data files
```{r}

## List with all xls files
all_xls <- tibble(list.files(raw_data_path, pattern = "animals")) %>% #c("xls", "xlsx")
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         ### UPDATE
         type = "animals") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)

```


## Read, explore, clean, save

### Data set 1: Intertidal Organisms

Data looks pretty clean, The main question is why are there two Amp (v) columns.
20 variables

**Date Range**
"2009-04-30 UTC" "2010-01-06 UTC"
```{r}
## intertidal organisms

## Explore
excel_sheets(all_xls$path[1])

## Read
intertidal_org <- read_excel(all_xls$ path[1], sheet = "Sheet1") %>% 
  clean_names()

## Clean
intertidal_clean <- intertidal_org %>% 
  rename(amp_1 = amp_v_13, ## v = volts
         amp_2 = amp_v_14,
         wt_n = wt_percent_n,
         wt_c = wt_percent_c)

## Save
write_csv(intertidal_clean, here::here("clean_data/2009_2010_palmyra_intertidal_organisms_isotope.csv"))

```


Data set 2: Rat data

Data is for the most part clean. Only has 14 variables and column name partially match the intertidal data.

**Date range**
2009-04-28 UTC" "2009-10-31 UTC"

```{r}
## Rat isotopes
excel_sheets(all_xls$path[2]) ## 2 sheets only sheet 1 has data

## Read
rat_data <- read_excel(all_xls$path[2], sheet = "Sheet1") %>% 
  clean_names()

## Clean
rat_clean <- rat_data %>% 
  rename(amp_1 = amp_v_6,
         amp_2 = amp_v_7,
         site = island_name,
         wt_n = wt_percent_n,
         wt_c = wt_percent_c) %>% 
  select(-x2)

## Save
write_csv(rat_clean, here::here("clean_data/2009_palmyra_rat_isotope.csv"))

```

Data set 3: Crab data

15 variables
two column with no name
similar measurements. Differences are in identification columns

**Date range**
2009-09-21 UTC" "2009-10-25 UTC

```{r}
## crab isotopes
excel_sheets(all_xls$path[3]) ## 2 sheets only sheet 1 has data

## Read
crab_data <- read_excel(all_xls$path[3], sheet = "Sheet1") %>% 
  clean_names()

## Clean
crab_clean <- crab_data %>% 
  rename(id = x2,
         site = island_name,
         amp_1 = amp_v_6,
         amp_2 = amp_v_7,
         wt_n = wt_percent_n,
         wt_c = wt_percent_c,
         organism_species = x15)

## Save
write_csv(crab_clean, here::here("clean_data/2009_palmyra_crab_isotope.csv"))

```

Data set 4: Amphipod and Isopod data

17 columns, last 2 have no relevant data
Two columns with no name
Similar measurements to all other data sets. Main differences is in how the recorded the ids for each measurement.
Date needs to be fixed!


```{r}
## Amphipod and isopod data
excel_sheets(all_xls$path[4]) ## 2 sheets only sheet 1 has data

## Read
amph_data <- read_excel(all_xls$path[4], sheet = "Sheet1") %>% 
  clean_names()

## Clean
amph_clean <- amph_data %>% 
  rename(id_1 = x3,
         id_2 = x4,
         id_3 = x6,
         site = island_name,
         amp_1 = amp_v_9,
         amp_2 = amp_v_10,
         wt_n = wt_percent_n,
         wt_c = wt_percent_c) %>% 
  mutate(date = case_when(date_run != "March 23 2010" ~ date_run),
         date = as.numeric(date),
         date = as.POSIXct(date*86400, origin = "1899-12-30", tz = "HAST"),
         date = case_when(is.na(date) ~ as.POSIXct(ymd("20100323"), tz = "GMT"),
                          T ~ date)) %>% 
  select(date, everything(), -x16, -x17, -date_run)
  
## SOLVE DATE ISSUE ##

## Save
write_csv(amph_clean, here::here("clean_data/"))

```

Data set 5: Insect data 2010

15 variables 
All column have names
Two Amp_v columns

```{r}
## insect data
excel_sheets(all_xls$path[5]) ## 2 sheets only sheet 1 has data

insect_data <- read_excel(all_xls$path[5], sheet = "Sheet1", skip = 4) %>% 
  clean_names()
```

Data set 6: Geograpsus Crab

Contains weight and length data for this species (Does this data belong to body size package?)
Here we are only going to clean isotope data

```{r}
## georapsus crab data
excel_sheets(all_xls$path[6]) ## only reading isotopes sheets. "isotopes 2013" & "isotopes 2014-2015"


gcrab_13 <- read_excel(all_xls$path[6], sheet = "isotopes 2013") %>% 
  clean_names()

gcrab_14_15 <- read_excel(all_xls$path[6], sheet = "isotopes 2014-2015") %>% 
  clean_names()


gcrab_13_clean <- gcrab_13 %>% 
  mutate(year = 2013) %>% 
  select(year, everything(), -d13c_comment, -estimated_enrichment)

gcrab_14_15_clean <- gcrab_14_15 %>% 
  rename(site = islet) %>% 
  select(year, everything())


```





