---
title: "Cleaning soil isotopes data"
author: "Camila Vargas"
date: "10/8/2021"
output: html_document
---

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)

## Sets R not to use scientific notations
options(scipen=999) 

```


## Check all necesary files are in the raw_data folder

```{r}
## File path to raw data used in this script
raw_data_path <- here::here("raw_data")

# Count files inside the raw_data folder to make sure the number of files downloaded is what is expected.
length(list.files(raw_data_path, pattern = "soil"))

## in this case there should be **4** files
```


## Table with all the raw data files
```{r}

## List with all xls files
all_xls <- tibble(list.files(raw_data_path, pattern = "soil")) %>% #c("xls", "xlsx")
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         ### UPDATE
         type = "soil") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)

```


## Read data

- 2007 and 2008 data
```{r}
## check sheet names
excel_sheets(all_xls$path[1]) ## Only sheet one has data

soil_07_08 <- read_excel(all_xls$path[1], sheet = "Sheet1") %>% 
  clean_names()

```

- 2011 data
```{r}
## check sheet names
excel_sheets(all_xls$path[2]) ## Only sheet one has data

soil_11 <- read_excel(all_xls$path[2], sheet = "Sheet1", skip = 4) %>% 
  clean_names()
```


- 2012 data
```{r}
## check sheet names
excel_sheets(all_xls$path[3]) ## Only sheet one has data

soil_12 <- read_excel(all_xls$path[3], sheet = "Sheet1", skip = 4) %>% 
  clean_names()

```

### Notes

Data set 11 and 12 all columns match. 11 and 12 are the same data sets!

There are slight differences in the decimals but technically they are the same dataset?

Functions to check if two data frames are de same:
`identical()`
`all_equal()`

**2012 Data**
id_number_1_1 = date
id_number_1_2 = island
id_number_2 = Location (Palmyra)
id_number_3 = type (soil)

Rename and add a column named horizon with NAs

**Data 07-08**
id_number_1 = number
id_number_2 = type (soil)
id_number_3 = Location (Palmyra)

Remove id_number 1, rename id_2 and id_3

## Organizing data so all columns match

```{r}

new_07_08 <- soil_07_08 %>% 
  rename(type = id_number_2,
         location = id_number_3,
         amp_v_n = amp_v_6,
         amp_v_c = amp_v_7) %>% 
  select(-id_number_1) %>% 
  mutate_at(vars(wt_percent_c, d13c, c_n), as.character)


new_2012 <- soil_12 %>% 
  filter(amp_v_6 != "N") %>% 
  rename(date = id_number_1_1,
         island = id_number_1_2,
         location = id_number_2,
         type = id_number_3,
         amp_v_n = amp_v_6,
         amp_v_c = amp_v_7) %>% 
  mutate(horizon = NA) %>% 
  mutate_at(vars(5:8,10), as.numeric)

```

## All data together

```{r}
soil_complete <- bind_rows(new_07_08, new_2012)
```

***Question for ANA***
Are the amp_v values on 07-08 data for N and C respectively on columns 8 and 9?

