---
title: "Cleaning palnt isotope data"
author: "Camila Vargas"
date: "10/6/2021"
output: html_document
---

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)

## Sets R not to use scientific notations
options(scipen=999) 

```


## Check all necesary files are in the raw_data folder

```{r}
## File path to raw data used in this script
raw_data_path <- here::here("raw_data")

# Count files inside the raw_data folder to make sure the number of files downloaded is what is expected.
length(list.files(raw_data_path, pattern = "plants"))

## in this case there should be **4** files
```


## Table with all the raw data files
```{r}

## List with all xls files
all_xls <- tibble(list.files(raw_data_path, pattern = "plants")) %>% #c("xls", "xlsx")
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         ### UPDATE
         type = "plant") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)

```

**Initial Scan**
- There are 4 files for this project. One for each of the years 2010, 2011, 2012 and one compile dataset with data from 2009 to 2012.

- Data for 2011 has information for only one plant

- Data for 2010 and 2012: is duplicate (with maybe extra) of 2010-2012 plant datasheet 


**Next Steps**
1. Read and clean compile data (2009 to 2010)
2. Identify how similar is the compile data with the datasets for each spcific year


## Read data

Compile file

**Note**
- Sheet2 contains short metadata information

- Sheet1 has the the actual data
```{r}
## check sheet names
excel_sheets(all_xls$path[1])

plant_compile <- read_excel(all_xls$path[1], sheet = "Sheet1") %>% 
  clean_names() %>% 
  rename(species_code = species_1,
         species = species_12,
         amp_n = amp_v_5,
         amp_c = amp_v_6) %>% 
  mutate(year = date)

```

Initial explorations
```{r}
str(plant_compile)
unique(plant_compile$date)

# Check data NAs
filter(plant_compile, is.na(date))

# How many observations by year?
plant_compile %>% 
  group_by(date) %>% 
  tally()
```



Data per year

- The 2011 data has one entry and columns not necessarily matches the compile dataset. Therefore we are not going to incorporate this data into the "clean data" outcome of this project. We are not sure what is the correct way to include this data. Please see 2011_Palmyra_Plant_Isotop.xlsx in [Miller-ter Kuile's Github account](https://github.com/anamtk/palmyra/tree/main/raw_data/isotope_data/plants) if you need the data on this one plant.


```{r}

# 2010
## check sheet names
excel_sheets(all_xls$path[2])

plant_2010 <- read_excel(all_xls$path[2], sheet = "Sheet1") %>% 
  clean_names() %>% 
  rename(species_code = id_number_1,
         island_name = id_number_2)
## Actually has data for 2011

# 2011
excel_sheets(all_xls$path[3])

plant_11 <- read_excel(all_xls$path[3], sheet = "Sheet1") %>% 
  clean_names()

## check sheet names
excel_sheets(all_xls$path[4])

plant_2012 <- read_excel(all_xls$path[4], sheet = "Sheet1", skip = 4) %>% 
  clean_names() %>% 
  rename(species_code = id_number_2,
         island_name = id_number_1)
## this has data from 2012

```


Exploring each of the data sets 
```{r}
colnames(plant_2010)
colnames(plant_11)
colnames(plant_2012)

```

**Notes**
- plant_2010 under data column there are only 2011 dates.
- plant_11 has one entry and date is on 2012
- All the columns names for these three data set match

## Combining yearly data

```{r}

  plant_all_yrs <- plant_11 %>% 
  rename(date = id_number_1_1,
         species_code = id_number_1_2,
         island_name = id_number_2) %>% 
  bind_rows(plant_2010, plant_2012) %>% 
  filter(d15n != "Air") %>% 
  rename(species = x14,
         amp_n = amp_v_6,
         amp_c = amp_v_7 ) %>% 
  select(-id_number_3, -run) %>% 
  mutate_at(vars(amp_n, amp_c, d15n, d13c, c_n), as.numeric) %>% 
  mutate(year = year(date))

```

## Compare both data sets

The compile against the all years

Both have the same measurements. The big question is: is the compile a subset of the all_yrs data?

```{r}

setdiff(plant_compile$wt_mg, plant_all_yrs$wt_mg)

setdiff(plant_all_yrs$wt_mg, plant_compile$wt_mg)

setdiff(plant_compile$amp_n, plant_all_yrs$amp_n)

setdiff(plant_all_yrs$amp_n, plant_compile$amp_n)

```

Looking at the difference between the two data in the different numeric parameters. There is no consistency on values that match between the two sets and that are different. 
This means that the compile is not a subset of the all year data.

Next step is to bind them together and remove duplicates.

## Row bind and remove duplicates

```{r}

all_data <- plant_compile %>% 
  mutate(date = NA) %>% 
  bind_rows(plant_all_yrs) %>% 
  select(-species_code, -date) %>% 
  mutate(island_name = recode(island_name,
                              "Cast" = "Castor",
                              "Dudl" = "Dudley",
                              "East" = "Eastern",
                              "Engi" = "Engineer",
                              "Frig" = "Frigate",
                              "Hole" = "Holei",
                              "Kaul" = "Kaula",
                              "Para" = "Paradise",
                              "Port" = "Portsmouth",
                              "Saci" = "Sacia"),
         type = "plant") %>%
  distinct()

```


For the most part ready! need to figure put how to add the date to the records that have Ana actual date!

And decide if we want to publish each data set separately or all together. This might be a question for Ana.





