---
title: "Download data from Drive"
author: "Camila Vargas"
date: "10/6/2021"
output: html_document
---

This script downloads the data needed for this project from a specific folder on our Google Drive

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googledrive)
library(here)
library(tidyverse)


## Sets R not to use scientific notations
options(scipen=999) 

```

## General Function
```{r}
download_data <- function(url,
                          code_name){
  folder_url <- url

# list of files inside the folder
files <- drive_ls(as_id(folder_url))


## Download all file to local computer. 
purrr::walk2(
    map(files$id, as_id),
    paste0("raw_data/", code_name,"_", files$name),
    drive_download,
    overwrite = TRUE)
                          }
```




## Download data files into local computer

Use the `download_data()` function to get the data from any folder in the drive.
url = url of folder you want to get the data
code_name = prefix you want you data set name yo have if any

1. Plants data

```{r download weather files}

download_data(url = "https://drive.google.com/drive/u/0/folders/1x5BfrEK2dUL1iaKznTUzOigpC1WqaDvG",
            code_name = "plants")

```

2. Spider

```{r}


spider_folder_url <- "https://drive.google.com/drive/u/0/folders/1QyMr6GOTCYCPIEAuO4ksf_LIgPFiq15Q"


download_data(url = spider_folder_url,
              code_name = "spider")

```


3. Soil

```{r}

# url of folder where the data lives
# Copy paste the url of the folder where the data lives

soil_folder_url <- "https://drive.google.com/drive/u/0/folders/1YbNq0k2_7GBDBh_U3-Md_sRJd4WlD4Lm"

download_data(url = soil_folder_url,
              code_name = "soil")
```

4. Baselines
```{r}

# url of folder where the data lives
# Copy paste the url of the folder where the data lives

baseline_folder_url <- "https://drive.google.com/drive/u/0/folders/1y1nixYCL8O1DgdryYCUTpTMCfYT4M8YU"

download_data(url = baseline_folder_url,
              code_name = "baseline")

```


5. Other Animals

```{r}
animals_folder_url <- "https://drive.google.com/drive/u/0/folders/1MDyF2fajJW3bUUhvtozq1vv5dZeGjg0k"

download_data(url = animals_folder_url,
              code_name = "animals")

```

6. Geckos
```{r}
geckos_folder <- "https://drive.google.com/drive/u/0/folders/1slotCD-6vBm2ZpFW1qXwcAD_0B_PdfoN"

download_data(url = geckos_folder,
              code_name = "geckos")
```

